---
title: "Correlation Coefficient"
author: "Claire Miller"
date: "11/07/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
# Libraries
library(malariaAtlas)
library(raster)
library(ggspatial)

# Files
hum_filename <- "ProjectHum/Monthly Outputs/AggScale.grd"
temp_filename <- "ProjectTemp/Monthly Outputs/AggScale.grd"
rain_filename <- "ProjectSurface/Monthly Outputs/Agg.grd"
# Layers
mal_rate_layer <- "Plasmodium falciparum PR2 - 10 version 2020"
# Use the africa layer to get the domain
layer_africa_2015 <- "Plasmodium falciparum PR2-10"

## Switch to just Africa but latest data

# Data processing functions
processEnvRaster <- function(raster_filename,label,extentRaster=mal_rate) {
  x <- raster(raster_filename)
  y <- crop(x, extent(extentRaster))
  names(y) = label
  return(y)
}
getRasterCorrelation <- function(x1,x2=mal_rate,conv_na=T) {
  x <- stack(x1,x2)
  x.df <- as.data.frame(x,xy=T)
  # Want to assign 0 to any NA malaria values that are on land
  if (conv_na) {
    x2.na <- which(is.na(x.df[,names(x2)]) & !is.na(x.df[,names(x1)]))
    x.df[x2.na,names(x2)] = 0
  }
  cor(x.df[,names(x1)],x.df[,names(x2)],
      method = "spearman", use="complete.obs")
}
```

Want to determine a correlation coefficient between a determined suitability index and the estimated malaria prevalence from the marlia Atlas.
Have three indices: humidity (`ProjectHum`), temperature (`ProjectTemp`), and rainfall (`ProjectSurface`).
Maps are available for the monthly data, but we only use the yearly aggreagated data (`AggScale.grd`) for each variable.
Any pixel that has no data (i.e. NA) on the malaria incidence raster, but does have data on the indices raster, is assumed to have a malaria incidence of 0. 

# Malaria data

Malaria data is downloaded from the Malaria Atlas project.
We use the layer: ``r mal_rate_layer`` with the bounding box for Africa taken from ``r layer_africa_2015``.
We only focus on Africa as my latop can't deal with the whole globe scale.

```{r mal-atlas, fig.width = 10, include=F}
africa_rate_2015 <- getRaster(surface=layer_africa_2015)
mal_rate <- getRaster(surface=mal_rate_layer, extent = bbox(africa_rate_2015))
#mal_rate.df <- as.MAPraster(mal_rate)
ggplot() + layer_spatial(mal_rate) +
  labs(fill=mal_rate_layer)
```


# Humidity

Starting with humidity.
We are using ``r hum_filename``.
We firstly crop the extent to match that of the malaria data.

```{r humidity-data}
hum <- processEnvRaster(hum_filename,"Humidity")
ggplot() + layer_spatial(hum) + labs(fill="Humidity")
```

Then we can correlate the humidity and the Malaria rate.
We just get the value at each pixel for both and run the base `cor` method.

```{r hum-cor}
cat("Spearman correlation between humidity and malaria rate: ",getRasterCorrelation(hum),"\n")
```

# Temperature

We are using ``r temp_filename``.
We firstly crop the extent to match that of the malaria data.

```{r temperature-data}
temp <- processEnvRaster(temp_filename,"Temperature")
ggplot() + layer_spatial(temp) + labs(fill="Temperature")
```

Then we can correlate the humidity and the Malaria rate.
We just get the value at each pixel for both and run the base `cor` method.

```{r temp-cor}
cat("Spearman correlation between temperature and malaria rate: ",getRasterCorrelation(temp),"\n")
```


# Rainfall

We are using ``r rain_filename``.
We firstly crop the extent to match that of the malaria data.

```{r rainfall-data}
rain <- processEnvRaster(rain_filename,"Rainfall")
ggplot() + layer_spatial(rain) + labs(fill="Rainfall")
```

Then we can correlate the humidity and the Malaria rate.
We just get the value at each pixel for both and run the base `cor` method.

```{r rain-cor}
cat("Spearman correlation between rainfall and malaria rate: ",getRasterCorrelation(rain),"\n")
```
